Probability

### 2016-04-13 prob.tex
prob

\documentclass{article}
\usepackage{fullpage}
\usepackage{amssymb}
\usepackage{amsthm}

\theoremstyle{remark}
\newtheorem{example}{Example}
\newcommand{\pmeas}{\ensuremath{\mathbb{P}}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Classical Conventional Probability Spaces}

Textbook probability theory is defined using the notions of
\emph{sample spaces}, \emph{events}, and
\emph{measures}~\cite{GrahamKnuthPatashnik1994,inun.425605319950101,rohatgi2011introduction}.

@@

@@

\ldots

>>>>

\documentclass{article}
\usepackage{fullpage}
\usepackage{bbm}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multicol}

\theoremstyle{remark}
\newtheorem{example}{Example}
\newcommand{\events}{\ensuremath{\mathcal{F}}}
\newcommand{\qevents}{\ensuremath{\mathcal{A}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\pr}[2]{\langle #1, #2 \rangle}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Probability Spaces}

%%%%%
\subsection{Classical Probability Spaces}

Textbook probability
theory~\cite{inun.425605319950101,GrahamKnuthPatashnik1994,rohatgi2011introduction}
is defined using the notions of a \emph{sample space} $\Omega$, a
space of \emph{events}~$\events$, and a \emph{probability
  measure}~$\pmeas$. In this paper, we will only consider
\emph{finite} sample spaces: we therefore define a sample space
$\Omega$ as an arbitrary non-empty finite set and the space of events
$\events$ as, $2^\Omega$, the powerset of $\Omega$. A \emph{probability
measure} is a function $\pmeas : \events \rightarrow [0,1]$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = 1$, and 
\item for a collection of pairwise disjoint events $E_i$, we have
  $\pmeas(\bigcup E_i) = \sum \pmeas(E_i)$. 
\end{itemize}

\begin{example}[Two coin experiment] Consider an experiment that
  tosses two coins. We have four possible outcomes that constitute the
  sample space $\Omega = \{ HH, HT, TH, TT \}$. The event that the
  first coin is ``heads'' is $\{ HH, HT \}$; the event that the two
  coins land on opposite sides is $\{ HT, TH \}$; the event that at
  least one coin is tails is $\{ HT, TH, TT\}$. Depending on the
  assumptions regarding the coins, we can define several probability
  measures. Here is a possible one:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& 0 \\
\pmeas(\{ HH \}) &=& 1/3 \\
\pmeas(\{ HT \}) &=& 0 \\
\pmeas(\{ TH \}) &=& 2/3 \\
\pmeas(\{ TT \}) &=& 0 \\
\pmeas(\{  HH, HT \}) &=& 1/3 \\
\pmeas(\{  HH, TH \}) &=& 1 \\
\pmeas(\{  HH , TT \}) &=& 1/3 
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& 2/3 \\
\pmeas(\{  HT , TT \}) &=& 0 \\
\pmeas(\{  TH , TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH \}) &=& 1 \\
\pmeas(\{  HH, HT, TT \}) &=& 1/3 \\
\pmeas(\{  HH, TH, TT \}) &=& 1 \\
\pmeas(\{  HT, TH, TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH, TT \}) &=& 1
\end{array}
\end{array}\]
\end{example}

%%%%%
\subsection{Quantum Probability Spaces}

A classical model decides the occurrence or non-occurence of all
events simultaneously which is inconsistent with quantum
mechanics. Indeed, in the quantum world, there are (non-commuting)
events which cannot happen simultaneously. To accommodate this
situation, we completely abandon the sample space $\Omega$ and define
and reason directly about events. Thus a quantum probability space
will consist of just two components: a set of events $\qevents$ and a
probability measure $\phi : \qevents \rightarrow [0,1]$. These
components are defined as follows.

We first assume an ambient Hilbert space $\Hilb$ and define the set of
events $\qevents$ as \emph{projections} on $\Hilb$. Similarly to the
classical case, a probability measure is a function
$\phi : \qevents \rightarrow [0,1]$ satisfying:
\begin{itemize}
\item $\phi(\mathbbm{1}) = 1$, and
\item for all $A \in \qevents$, we have $\phi(A^*A) \geq 0$.
\end{itemize}

As an example, let $P_1, P_2, \ldots, P_k$ be mutually orthogonal
projections on $\Hilb$ with sum $\mathbbm{1}$ and define the event
space $\qevents$ to be the linear span of these operators:
\[
\qevents = \{ \sum_{j=1}^k \lambda_jP_j ~|~ \lambda_i, \ldots, \lambda_k \in \mathbb{C} \}. 
\]
Each state $\psi$ of the Hilbert space induces a probability measure
$\phi_\psi : \qevents \rightarrow [0,1]$ defined as follows:
\[
\phi_\psi(A) = \langle \psi ~|~ A\psi \rangle
\]

%%%%%
\subsection{Plan}

We see that a quantum probability space embeds several infinite
precision assumptions, both in the Hilbert space itself and in the
range of the probability measure. We will now study the consequences
of restricting each occurrence of $\mathbb{R}$ or $\mathbb{C}$ to some
finite structure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We begin by reviewing the conventional presentation of classical
probability spaces and then give an alternative formulation that is
``quantum-like'' but still classical. We conclude this section with a
definition of quantum probability spaces given as a modest
generalization of the alternative classical definition.

%%%%%
\subsection{Conventional Classical Probability Spaces}

Textbook probability
theory~\cite{inun.425605319950101,GrahamKnuthPatashnik1994,rohatgi2011introduction}
is defined using the notions of a \emph{sample space} $\Omega$, a
space of \emph{events}~$\events$, and a \emph{probability
  measure}~$\pmeas$. In this paper, we will only consider
\emph{finite} sample spaces: we therefore define a sample space
$\Omega$ as an arbitrary non-empty finite set and the space of events
$\events$ as, $2^\Omega$, the powerset of $\Omega$. A \emph{probability
measure} is a function $\pmeas : \events \rightarrow [0,1]$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = 1$, and 
\item for a collection of pairwise disjoint events $E_i$, we have
  $\pmeas(\bigcup E_i) = \sum \pmeas(E_i)$. 
\end{itemize}

\begin{example}[Two coin experiment] Consider an experiment that
  tosses two coins. We have four possible outcomes that constitute the
  sample space $\Omega = \{ HH, HT, TH, TT \}$. The event that the
  first coin is ``heads'' is $\{ HH, HT \}$; the event that the two
  coins land on opposite sides is $\{ HT, TH \}$; the event that at
  least one coin is tails is $\{ HT, TH, TT\}$. Depending on the
  assumptions regarding the coins, we can define several probability
  measures. Here is a possible one:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& 0 \\
\pmeas(\{ HH \}) &=& 1/3 \\
\pmeas(\{ HT \}) &=& 0 \\
\pmeas(\{ TH \}) &=& 2/3 \\
\pmeas(\{ TT \}) &=& 0 \\
\pmeas(\{  HH, HT \}) &=& 1/3 \\
\pmeas(\{  HH, TH \}) &=& 1 \\
\pmeas(\{  HH , TT \}) &=& 1/3 
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& 2/3 \\
\pmeas(\{  HT , TT \}) &=& 0 \\
\pmeas(\{  TH , TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH \}) &=& 1 \\
\pmeas(\{  HH, HT, TT \}) &=& 1/3 \\
\pmeas(\{  HH, TH, TT \}) &=& 1 \\
\pmeas(\{  HT, TH, TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH, TT \}) &=& 1
\end{array}
\end{array}\]
\end{example}

%%%%%
\subsection{Alternative Definition of Classical Probability Spaces}

In the conventional presentation, we have viewed the space of events
$2^\Omega$ are the powerset of $\Omega$. We can equivalently view
$2^\Omega$ as the space of functions from $\Omega$ to the set
$2 = \{0,1\}$. For example, the event $\{HT,TH\}$ is the function $e$
such that:
\[
e (HH) = 0, \quad e (HT) = 1, \quad e (TH) = 1, \quad e (TT) = 0 
\]
We will in fact do a sweeping generalization and view events as
functions from $\Omega$ to $\mathbb{C}$, the set of complex
numbers. This accommodates the previous events such as $e$ and allows
many more events such as event $e'$ below:
\[
e' (HH) = \sqrt{2} + i \sqrt{3}, \quad e' (HT) = 1, \quad e' (TH) = \pi, \quad e' (TT) = 0 
\]
The events are not going to be completely arbitrary functions,
however. We will insist on some conditions:...

This generalization

@@
